---
title: "RegressionFinalProject"
author: "MK"
date: "November 12, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
#Excutive Summary:
In this regression analysis, first a fit between mpg and regressors is calculated. In this fit, the am variable (transmission kind) does not have significant contribution based on the P value obtained from anova function. Another fit is introduced which indicates relationship between transmission and mpg. The results show that manual transmission is better for MPG.

# Analysis:
In this analysis the contribution of every regressor in the linear regression that predicts mpg is studied by evaluating their P values of nested models. We first start with a simple regression between mpg and cyl.

```{r}
lmd1 <- lm(mpg~factor(cyl), data = mtcars)
summary(lmd1)$coef
```

Now, we add other regressors to this model one by one and anylize their necessity via anova function.
```{r}
lmd2 <- lm(mpg~factor(cyl)+disp-1, data = mtcars)
anova(lmd1, lmd2)
```
The resultant p value indicates that the new term in mdl2 is neccessary over mdl1.

We add a new term which is the interaction between disp and factor(cyl).
```{r}
lmd3 <- lm(mpg~factor(cyl)+disp+disp*factor(cyl), data = mtcars)
anova(lmd1, lmd2, lmd3)
```
As p value indicates, the new term in mdl3 is necessary over mdl2.

Let's take a look at the fittings.
```{r, fig.width = 4, fig.height = 4}
plot(mtcars$disp, mtcars$mpg, type = 'n', frame=FALSE, xlab = 'disp', ylab = 'mpg')
abline(lm(mtcars$mpg~mtcars$disp), lwd = 2)
abline(coef(lmd3)[1], coef(lmd3)[4], col="red")
points(mtcars$disp[mtcars$cyl==4], mtcars$mpg[mtcars$cyl==4],
       pch=21, col='black', bg="red", cex=2)
abline(coef(lmd3)[1]+coef(lmd3)[2], coef(lmd3)[4]+coef(lmd3)[5], col='blue')
points(mtcars$disp[mtcars$cyl==6], mtcars$mpg[mtcars$cyl==6],
       pch=21, col='black', bg="blue", cex=2)
abline(coef(lmd3)[1]+coef(lmd3)[3], coef(lmd3)[4]+coef(lmd3)[6], col='dark green')
points(mtcars$disp[mtcars$cyl==8], mtcars$mpg[mtcars$cyl==8],
       pch=21, col='black', bg="dark green", cex=2)
legend('topright', legend = c('all together', '4 cyl','6 cyl','8 cyl'),
       col = c('black','red', 'blue', 'dark green'), lty=1,
       lwd = 2)
```

As shown in the graph, the four cylinder cars mpg has stronger relationship with disp regressor.

If we add other variables(regressors) one by one and study their significance using anova function, we will see that none of the remaining variables (including transmission(am)) are significant for the regression model.

So far, it was shown that lmd3 is the best model for mtcar data that shows the relationship between mpg and other variables. We can further analyze lmd3 as follows:
```{r, fig.width = 3, fig.height = 3}
plot(lmd3)
```

The Residuals vs Fitted plot indicates there is no relationship between Residuals and Fitted values which shows a good fit. The Q-Q plot shows that the residuals distribution is normal. And the Standardized Residulas plots show the presence of one high leverage point (Hornet 4 Drive) which does not seem to be influential.

Based on this analysis the am (transmission kind) variable is not necessary for mpg prediction if we use the variables in lmd3.

However, to better understand the relationship between mpg and transmission variable, we can start our model with using the am variable.
```{r}
lmd12 <- lm(mpg~factor(am), data = mtcars)
summary(lmd12)$coef
```
Based on the P values, it can be concluded that there is a relationship between transmission kind and mpg.

The lmd12 coefficients show that the average of mpg increases by around seven units for manual transmission with refrence to automatic transmission with an mpg average of around 17.

This model can be further studied by adding other variables to the model and applying anova analysis to see if the added variable is necessary or not.

Lets add the cyl variable.
```{r}
lmd13 <- lm(mpg~factor(am) + factor(cyl) + factor(cyl)*factor(am), data = mtcars)
anova(lmd12, lmd13)
```
The anova analysis results indicate that the cyl term is necessary for mpg analysis.

To better understand the relationship between the am and cyl variables, we should study the fitting coefficients of lmd13.
```{r}
summary(lmd13)$coef
```
The coefficients show that switching from automatic to manual transmission will increase the average of mpg.

This fact can be visualized as follows:
```{r, fig.width = 10, fig.height = 4}
require(ggplot2)
transmission <- mtcars$am
transmission[mtcars$am == 0] <- 'automatic'
transmission[mtcars$am == 1] <- 'manual'
temp <- interaction( transmission, factor(mtcars$cyl) )
ggplot( aes(y=mpg, x=temp, fill=factor(mtcars$cyl)), data = mtcars ) + geom_boxplot()
```

